Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-06-18 forecast/paper/paper.tex
fixes to figure sizes / margins

% Please give the surname of the lead author for the running footer
\leadauthor{Lead author last name}

@@

Figure~\ref{fig:forecast} displays the results for forecasting, stratified by
ahead value and averaged over all HRRs and forecast dates (June 9--December 31,
2020).  Shown is the average WIS for each forecast model divided by the averaged
WIS from a baseline model, which is basically a flat-line forecaster (its
median forecast for $Y_{\ell,t+a}$ is always $Y_{\ell,t}$, with predicted
quantiles defined around this flat line based on historical variation). This is
the same baseline model as that used in the COVID-19 Forecast Hub.  Here, we  
divide by the average WIS of the baseline model in order to put the y-axis on an 
interpretable scale.  We can see that all curves are below 1, which means
(smaller WIS is better) that all of the models, including AR, outperform the
baseline on average over the forecasting period.  On the other hand, the models 
deliver at best an improvement of about 20\% in mean WIS over the baseline
model, with this gap narrowing to only 10\% at the largest ahead values,
illustrating the difficulty of the forecasting problem.

@@

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{fig/fcast-1.pdf}
  \caption{Average WIS for each forecast model, over all forecast dates and all
    HRRs, divided by the average WIS achieved by a baseline model (a
    probabilistic version of the flat-line forecaster).  Here and in all figures
  we abbreviate CTIS-CLI-in-community by CTIS-CLIIC.}  
  \label{fig:forecast}
\end{figure}

Figure~\ref{fig:hotspot} displays the results for hotspot prediction, again
stratified by ahead value and averaged over all HRRs and prediction dates (June 
16--December 31, 2020).  We can see many similarities to the forecasting
results (now larger AUC is better); for example, CHNG-COVID and DV-CLI offer the
biggest improvement over AR, and all models including AR degrade in performance 
towards the baseline (this is now just a classifier based on random guessing,
which achieves an AUC of 0.5) as the ahead values grow.  A clear difference,
however, is that the CHNG-CLI model performs quite well in hotspot prediction,
close to the best-performing indicator models for many of the ahead values. This
may be because volatility in the CHNG-CLI indicator plays less of a role in the
associated logistic model's predicted probabilities (in general, a sigmoid
function can absorb a lot of the variability in its input).

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{fig/hot-1.pdf}
\caption{Area under the curve for each hotspot prediction model, calculated over
  all prediction dates and all HRRs.} 
\label{fig:hotspot}
\end{figure}

>>>>

% Please give the surname of the lead author for the running footer
\leadauthor{McDonald}

@@

Figure~\ref{fig:forecast} (right panel) displays the results for forecasting, stratified by
ahead value and averaged over all HRRs and forecast dates (June 9--December 31,
2020).  Shown is the average WIS for each forecast model divided by the averaged
WIS from a baseline model, which is basically a flat-line forecaster (its
median forecast for $Y_{\ell,t+a}$ is always $Y_{\ell,t}$, with predicted
quantiles defined around this flat line based on historical variation). This is
the same baseline model as that used in the COVID-19 Forecast Hub.  Here, we  
divide by the average WIS of the baseline model in order to put the y-axis on an 
interpretable scale.  We can see that all curves are below 1, which means
(smaller WIS is better) that all of the models, including AR, outperform the
baseline on average over the forecasting period.  On the other hand, the models 
deliver at best an improvement of about 20\% in mean WIS over the baseline
model, with this gap narrowing to only 10\% at the largest ahead values,
illustrating the difficulty of the forecasting problem.

@@

% \begin{figure}[t]
%   \includegraphics[width=\columnwidth]{fig/fcast-1.pdf}
%   \caption{Average WIS for each forecast model, over all forecast dates and all
%     HRRs, divided by the average WIS achieved by a baseline model (a
%     probabilistic version of the flat-line forecaster).  Here and in all figures
%   we abbreviate CTIS-CLI-in-community by CTIS-CLIIC.}  
%   \label{fig:forecast}
% \end{figure}

Figure~\ref{fig:forecast} (left panel) displays the results for hotspot prediction, again
stratified by ahead value and averaged over all HRRs and prediction dates (June 
16--December 31, 2020).  We can see many similarities to the forecasting
results (now larger AUC is better); for example, CHNG-COVID and DV-CLI offer the
biggest improvement over AR, and all models including AR degrade in performance 
towards the baseline (this is now just a classifier based on random guessing,
which achieves an AUC of 0.5) as the ahead values grow.  A clear difference,
however, is that the CHNG-CLI model performs quite well in hotspot prediction,
close to the best-performing indicator models for many of the ahead values. This
may be because volatility in the CHNG-CLI indicator plays less of a role in the
associated logistic model's predicted probabilities (in general, a sigmoid
function can absorb a lot of the variability in its input).

\begin{figure*}[t]
  \includegraphics[width=\textwidth]{fig/fcast-hot-combo-1.pdf}
  \caption{Left: Average WIS for each forecast model, over all forecast dates and all
    HRRs, divided by the average WIS achieved by a baseline model (a
    probabilistic version of the flat-line forecaster).  Here and in all figures
    we abbreviate CTIS-CLI-in-community by CTIS-CLIIC.
    Right: Area under the curve for each hotspot prediction model, calculated over
    all prediction dates and all HRRs.
  }  
  \label{fig:forecast}
\end{figure*}

@@

% \begin{figure}[t]
% \includegraphics[width=\columnwidth]{fig/hot-1.pdf}
% \caption{Area under the curve for each hotspot prediction model, calculated over
%   all prediction dates and all HRRs.} 
% \label{fig:hotspot}
% \end{figure}

### 2021-06-18 forecast/paper/supplement.Rmd
fixes to figure sizes / margins

```{r upswing-histogram, include=FALSE, fig.width=12}
up_down_df %>% 
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  ggplot(aes(AR_wis - wis, fill = forecaster)) +
  geom_histogram(bins = 100) +
  facet_grid(udf ~ forecaster) +
  theme_bw() +
  ylab("Count") +
  theme(legend.position = "none") +
  scale_fill_manual(values = fcast_colors) +
  scale_y_log10(breaks = c(1,100,10000), 
                labels = trans_format("log10", math_format(10^.x))) +
  coord_cartesian(ylim=c(1,2e4)) +
  xlab("AR WIS - forecaster WIS") +
  geom_vline(xintercept = 0)
```

@@

```{r leading-only, include=FALSE}
# Figure 6 in the manuscript
df2 %>%
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  summarise(
    leadingness = cor(AR_wis - wis, leading, use = "pairwise.complete.obs"),
    laggingness = cor(AR_wis - wis, lagging, use = "pairwise.complete.obs")) %>%
  ggplot(aes(udf, leadingness, fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_fill_manual(values = fcast_colors) +
  theme_bw() +
  ylab("Correlation") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

>>>>

```{r upswing-histogram, include=FALSE, fig.width=8, fig.height=4}
up_down_df %>% 
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  ggplot(aes(AR_wis - wis, fill = forecaster)) +
  geom_histogram(bins = 100) +
  facet_grid(udf ~ forecaster) +
  theme_bw() +
  ylab("Count") +
  theme(legend.position = "none") +
  scale_fill_manual(values = fcast_colors) +
  scale_y_log10(breaks = c(1,100,10000), 
                labels = trans_format("log10", math_format(10^.x))) +
  coord_cartesian(ylim=c(1,2e4)) +
  xlab("AR WIS - forecaster WIS") +
  geom_vline(xintercept = 0)
```

@@

```{r leading-only, include=FALSE, fig.width=6, fig.height=4}
# Figure 6 in the manuscript
df2 %>%
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  summarise(
    leadingness = cor(AR_wis - wis, leading, use = "pairwise.complete.obs"),
    laggingness = cor(AR_wis - wis, lagging, use = "pairwise.complete.obs")) %>%
  ggplot(aes(udf, leadingness, fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_fill_manual(values = fcast_colors) +
  guides(fill = guide_legend(nrow = 1)) +
  theme_bw(base_size = 16) +
  ylab("Correlation") +
  xlab("") +
  theme(legend.position = c(1,-.15), 
        legend.justification = "right",
        legend.title = element_blank(), 
        legend.text = element_text(size = 10),
        plot.margin = margin(b = 20, l = 4),
        legend.background = element_rect(fill = "transparent"))
```
